---
title: "Behavioral Experiment Analyses"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-expand: 1
    toc-title: Contents
    toc-location: left
editor: visual
execute:
  freeze: auto
  echo: true
  message: false
  warning: false
---

# Load Libraries

```{r}
library(tidyverse)
library(knitr)
library(psych)
library(lmerTest) 
library(marginaleffects)
library(ordinal) 
library(emmeans)
```

# Load Data

## Participant-Level Data

```{r}
p_data = read_csv("participant_level_data.csv", show_col_types = FALSE)

kable(head(p_data), digits = 2, format = "markdown")
```

## Headline-Level Data

```{r}
head_data = read_csv("headline_level_data.csv", show_col_types = FALSE)

kable(head(head_data), digits = 2, format = "markdown")
```

# Annotator Agreement

## Annotators per Headline

```{r}
block_n = p_data |>
  distinct(wave, block, participant_id) |>
  count(wave, block, name = "n_participants")

knitr::kable(
  block_n |>
    summarise(
      `N blocks` = n(),
      `Mean annotators`    = mean(n_participants),
      SD       = sd(n_participants),
      Min      = min(n_participants),
      Max      = max(n_participants)
  ), digits = 2)
```

## Define Function

```{r}
# function to compute ICCs for a single block
compute_icc_per_block = function(block_data, rating_var) {
  wide_data = block_data %>%
    select(headline, participant_id, all_of(rating_var)) %>%
    pivot_wider(names_from = participant_id, values_from = all_of(rating_var))
  # convert to a matrix, excluding headline information
  rating_matrix = as.matrix(select(wide_data, -headline))
  # compute ICC, handle cases with too few raters
  icc_out = psych::ICC(rating_matrix)
  return(icc_out)
}
```

## Prepare Data

```{r}
# filter for clear headlines
# ensure villain and victim annotations are numeric
# create an identifier for each wave-block pair
p_data_agree = p_data %>%
  filter(clarity_binary == 1) %>%
  mutate(
    villain_presence_ordinal = as.numeric(villain_presence_ordinal),
    victim_presence_ordinal = as.numeric(victim_presence_ordinal),
    wave_block = paste0("W", wave, "_B", block)  # e.g., W1_B1, W2_B1
  )
```

```{r}
blocks = p_data_agree |>
  group_by(wave_block) |>
  group_split()

names(blocks) = p_data_agree |>
  group_by(wave_block) |>
  group_keys() |>
  pull(wave_block) |>
  as.character()
```

## Calculate

### Villain Ratings

```{r}
icc_villain = lapply(blocks, compute_icc_per_block,
                     rating_var = "villain_presence_ordinal")
```

```{r}
icc2k_list = list()

for (i in seq_along(icc_villain)) {
  
  icc = icc_villain[[i]]
  block_name = paste0("Block_", i)
  res = icc$results
  row = res["Average_random_raters", c("type", "ICC", "lower bound", "upper bound", "p")]
  icc2k_list[[i]] = data.frame(
    wave_block = block_name,
    type = row[["type"]],
    ICC2k = as.numeric(row[["ICC"]]),
    ICC2k_lower = as.numeric(row[["lower bound"]]),
    ICC2k_upper = as.numeric(row[["upper bound"]]),
    p = as.numeric(row[["p"]])
  )
}

# combine into a single dataframe
icc_summary = do.call(rbind, icc2k_list)

knitr::kable(icc_summary, digits = 2)
```

```{r}
round(mean(icc_summary$ICC2k), 2)
```

### Victim Ratings

```{r}
icc_victim  = lapply(blocks, compute_icc_per_block,
                     rating_var = "victim_presence_ordinal")
```

```{r}
icc2k_list = list()

for (i in seq_along(icc_victim)) {
  
  icc = icc_victim[[i]]
  block_name = paste0("Block_", i)
  res = icc$results
  row = res["Average_random_raters", c("type", "ICC", "lower bound", "upper bound", "p")]
  icc2k_list[[i]] = data.frame(
    wave_block = block_name,
    type = row[["type"]],
    ICC2k = as.numeric(row[["ICC"]]),
    ICC2k_lower = as.numeric(row[["lower bound"]]),
    ICC2k_upper = as.numeric(row[["upper bound"]]),
    p = as.numeric(row[["p"]])
  )
}

# combine into a single dataframe
icc_summary = do.call(rbind, icc2k_list)

knitr::kable(icc_summary, digits = 2)
```

```{r}
round(mean(icc_summary$ICC2k), 2)
```

# **Low-Quality News Sources are More Likely to Portray Villains and Victims**

## Prepare Data

### Participant-Level Data

```{r}
# code data types
p_data = p_data |>
  mutate(
    participant_id = as.factor(participant_id),
    headline_id    = as.factor(headline_id),
    gender         = as.factor(gender),
    party_id       = relevel(as.factor(party_id), ref = "independent"),
    across(c(ideology_ordinal, ladder_response, age, education_ordinal,
             wellbeing_ordinal, poli_trust_ordinal, moral_essentialism,
             affective_polarization, income), as.numeric),
    across(c(villain_presence_ordinal, victim_presence_ordinal),
           \(x) factor(x, ordered = TRUE))
  )

# code dyad variables
p_data = p_data |>
  mutate(
    villain_presence_binary = as.integer(villain_presence_binary),
    victim_presence_binary  = as.integer(victim_presence_binary),
    dyad_binary  = as.integer(villain_presence_binary == 1 &
                                victim_presence_binary  == 1),
    pc1_reversed = 1 - pc1,          # 0 = highest quality, 1 = lowest
    income_10k   = income / 10000,
    age_decade   = age / 10
  )

# subset on clear headlines
p_data_clear = p_data |> filter(clarity_binary == 1)

# grand-mean center covariates
covars_num = c("affective_polarization", "poli_trust_ordinal", "ideology_ordinal",
                "age_decade", "ladder_response", "income_10k",
                "education_ordinal", "moral_essentialism", "wellbeing_ordinal")
covars_all = c(covars_num, "gender", "party_id")

p_means = p_data_clear |>
  filter(complete.cases(pick(all_of(covars_all)))) |>
  distinct(participant_id, .keep_all = TRUE) |>
  summarise(across(all_of(covars_num), \(x) mean(x, na.rm = TRUE)))

p_data_clear = p_data_clear |>
  mutate(across(all_of(covars_num),
                \(x) x - p_means[[cur_column()]],
                .names = "{.col}_c"))
```

```{r}
# compute Ns
n_distinct(p_data_clear$headline_id)   
```

### Headline-Level Data

```{r}
head_data = head_data |>
  mutate(
    # code villains, victims, and the dyad
    villain_binary = as.integer(villain_presence_majority),
    victim_binary  = as.integer(victim_presence_majority),
    dyad_binary    = as.integer(villain_binary == 1 & victim_binary == 1),
    # reverse code source quality
    pc1_reversed = 1 - pc1
  )
```

```{r}
# subset on headlines rated as clear
head_data_clear = head_data %>%
  filter(clarity_majority == 1)

n_distinct(head_data_clear$headline_id)
```

## Analyses

### Participant-Level

#### Dyad \~ Source Quality

```{r}
m_dyad = glmer(
  dyad_binary ~
    pc1_reversed + 
    affective_polarization_c + poli_trust_ordinal_c + ideology_ordinal_c + party_id + 
    age_decade_c + gender + ladder_response_c + income_10k_c + education_ordinal_c +
    moral_essentialism_c + wellbeing_ordinal_c +
    (1 | participant_id) + (1 | headline_id),
  data    = p_data_clear,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
```

```{r}
kable(broom.mixed::tidy(
  m_dyad,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE,
  conf.method = "Wald"),
  digits = 4)
```

```{r}
# risk difference across the full source-quality range
avg_comparisons(m_dyad,
                variables = list(pc1_reversed = c(0, 1)),
                re.form = NA)
```

```{r}
# missingness
covars = c("affective_polarization", "poli_trust_ordinal", "ideology_ordinal",
           "party_id", "age_decade", "gender", "ladder_response", "income_10k",
           "education_ordinal", "moral_essentialism", "wellbeing_ordinal")

p_data_clear |>
  distinct(participant_id, .keep_all = TRUE) |>
  summarise(across(all_of(covars), ~ sum(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to = "Covariate", values_to = "N Missing") |>
  arrange(desc(`N Missing`)) |>
  mutate(
    Covariate = recode(Covariate,
                  "poli_trust_ordinal" = "Political trust",
                  "party_id" = "Political party",
                  "wellbeing_ordinal" = "Wellbeing",
                  "gender" = "Gender",
                  "education_ordinal" = "Education",
                  "income_10k" = "Income",
                  "affective_polarization" = "Affective polarization",
                  "ideology_ordinal" = "Political ideology",
                  "age_decade" = "Age",
                  "ladder_response" = "Subjective SES"
    )
  )
```

#### Villain \~ Source Quality

```{r}
m_villain = clmm(
  villain_presence_ordinal ~
    pc1_reversed + 
    affective_polarization_c + poli_trust_ordinal_c + ideology_ordinal_c + party_id + 
    age_decade_c + gender + ladder_response_c + income_10k_c + education_ordinal_c +
    moral_essentialism_c + wellbeing_ordinal_c +
    (1 + pc1_reversed | participant_id) + (1 | headline_id),
  data    = p_data_clear
)
```

```{r}
kable(broom.mixed::tidy(
  m_villain,
  conf.int = TRUE,
  exponentiate = TRUE), # exponentiate the results to odds ratios
  digits = 4)
```

```{r}
em = emmeans(m_villain, ~ pc1_reversed | villain_presence_ordinal,
              at = list(pc1_reversed = c(0, 1)), mode = "prob")
em
contrast(em, method = "revpairwise")
```

#### Victim \~ Source Quality

```{r}
m_victim = clmm(
  victim_presence_ordinal ~
    pc1_reversed + 
    affective_polarization_c + poli_trust_ordinal_c + ideology_ordinal_c + party_id + 
    age_decade_c + gender + ladder_response_c + income_10k_c + education_ordinal_c +
    moral_essentialism_c + wellbeing_ordinal_c +
    (1 + pc1_reversed | participant_id) + (1 | headline_id),
  data    = p_data_clear
)
```

```{r}
kable(broom.mixed::tidy(
  m_victim,
  conf.int = TRUE,
  exponentiate = TRUE), # exponentiate the results to odds ratios
  digits = 4)
```

```{r}
em = emmeans(m_victim, ~ pc1_reversed | victim_presence_ordinal,
              at = list(pc1_reversed = c(0, 1)), mode = "prob")
em
contrast(em, method = "revpairwise")
```

### Headline-Level

#### Dyad \~ Source Quality

```{r}
m_dyad_head = glmer(
  dyad_binary ~
    pc1_reversed +
    (1 | domain),
  data    = head_data_clear,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
```

```{r}
kable(broom.mixed::tidy(
  m_dyad_head,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE,
  conf.method = "Wald"),
  digits = 4)
```

#### Villain \~ Source Quality

```{r}
m_villain_head = glmer(
  villain_binary ~
    pc1_reversed +
    (1 | domain),
  data    = head_data_clear,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
```

```{r}
kable(broom.mixed::tidy(
  m_villain_head,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE,
  conf.method = "Wald"),
  digits = 4)
```

#### Victim \~ Source Quality

```{r}
m_victim_head = glmer(
  victim_binary ~
    pc1_reversed +
    (1 | domain),
  data    = head_data_clear,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
```

```{r}
kable(broom.mixed::tidy(
  m_victim_head,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE,
  conf.method = "Wald"),
  digits = 4)
```

### Robustness Analyses

Participant-level focal predictor models including all participants.

#### Dyad \~ Source Quality

```{r}
m_dyad_incl = glmer(
  dyad_binary ~
    pc1_reversed +
    (1 | participant_id) + (1 | headline_id),
  data    = p_data_clear,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
```

```{r}
kable(broom.mixed::tidy(
  m_dyad_incl,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE,
  conf.method = "Wald"),
  digits = 4)
```

#### Villain \~ Source Quality

```{r}
m_villain_incl = clmm(
  villain_presence_ordinal ~
    pc1_reversed +
    (1 + pc1_reversed | participant_id) + (1 | headline_id),
  data    = p_data_clear
)
```

```{r}
kable(broom.mixed::tidy(
  m_villain_incl,
  conf.int = TRUE,
  exponentiate = TRUE), # exponentiate the results to odds ratios
  digits = 4)
```

#### Victim \~ Source Quality

```{r}
m_victim_incl = clmm(
  victim_presence_ordinal ~
    pc1_reversed + 
    (1 + pc1_reversed | participant_id) + (1 | headline_id),
  data    = p_data_clear
)
```

```{r}
kable(broom.mixed::tidy(
  m_victim_incl,
  conf.int = TRUE,
  exponentiate = TRUE), # exponentiate the results to odds ratios
  digits = 4)
```

# **Dyadic Completion Helps Explain the Relative ‘Success’ of Villains**

## Prepare Data

```{r}
p_data_comp = p_data %>%
  # remove unclear headlines
  filter(clarity_binary == 1) %>%
  # clean free text fields
  mutate(
    # remove any whitespaces at the start or end and make all text lowercase
    victim_text_clean = str_trim(str_to_lower(victim_identity)),
    villain_text_clean = str_trim(str_to_lower(villain_identity)),
    # remove, nulls, blanks, and 'none'
    victim_text_clean = ifelse(
      victim_text_clean == "" |
        is.na(victim_text_clean) |
        str_detect(victim_text_clean, "^none$|^no$|^n/a$|^nil$"),
      NA, victim_text_clean
    ),
    villain_text_clean = ifelse(
      villain_text_clean == "" |
        is.na(villain_text_clean) |
        str_detect(villain_text_clean, "^none$|^no$|^n/a$|^nil$"),
      NA, villain_text_clean
    )
  ) %>%
  # binary indicators of whether an entity was named
  mutate(
    named_victim  = ifelse(is.na(victim_text_clean), 0, 1),
    named_villain = ifelse(is.na(villain_text_clean), 0, 1),
    # binary victim and villain presense on the likert measures presence 
    victim_present_bin  = ifelse(victim_presence_ordinal >= 4, 1, 0),
    villain_present_bin = ifelse(villain_presence_ordinal >= 4, 1, 0)
  )
```

```{r}
# aggregate to the headline level
head_data_comp = p_data_comp %>%
  group_by(headline_id) %>%
  summarise(
    villain_present_rate = mean(villain_present_bin, na.rm = TRUE),
    victim_present_rate  = mean(victim_present_bin, na.rm = TRUE),
    # completion 
    victim_named_rate  = mean(named_victim, na.rm = TRUE),
    villain_named_rate = mean(named_villain, na.rm = TRUE),
    # preserve source quality
    pc1 = mean(pc1, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )
```

```{r}
# define categories
head_data_comp = head_data_comp %>%
  mutate(
    dyad = case_when(
      villain_present_rate >= 0.5 & victim_present_rate < 0.5 ~ "villain_only",
      villain_present_rate < 0.5 & victim_present_rate >= 0.5 ~ "victim_only",
      villain_present_rate >= 0.5 & victim_present_rate >= 0.5 ~ "dyad",
      villain_present_rate < 0.5 & victim_present_rate < 0.5 ~ "neither",
      TRUE ~ NA_character_
    )
  )
```

```{r}
data_comp = head_data_comp %>%
  # select on cases with only a victim or only a villain
  filter(dyad %in% c("villain_only", "victim_only")) %>%
  # calculate rates of a victim being named in a villain-only case and
  # rates of a villain being named in a victim-only case
  mutate(
    completion_rate = case_when(
      dyad == "villain_only" ~ victim_named_rate,
      dyad == "victim_only"  ~ villain_named_rate
    ),
    completion_type = case_when(
      dyad == "villain_only" ~ "Victim completion",
      dyad == "victim_only"  ~ "Villain completion"
    )
  )
```

## Descriptives

```{r}
knitr::kable(data_comp %>%
  group_by(completion_type) %>%
  summarise(
    mean_completion = mean(completion_rate, na.rm = TRUE),
    sd_completion = sd(completion_rate, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ), digits = 2)
```

## Analysis

### t-Test

```{r}
t.test(
  completion_rate ~ completion_type,
  data = data_comp
)
```

# Session Information

```{r}
sessionInfo()
```
