---
title: "Probability Calibration"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 1
    toc-title: Contents
    toc-location: left
editor: visual
execute:
  freeze: auto
  echo: true
  message: false
  warning: false
---

Fits source-quality-aware recalibrations of the fine-tuned villain and victim classifiers using the held-out annotated test split

# Load Libraries

```{r}
library(tidyverse)
library(knitr)
library(broom)
library(rsample)
```

# Reconstruct the Annotated Test Split

```{r}
# train-test split record
manifest = read_csv(
  'train_test_split_record.csv',
  show_col_types = FALSE
  )

# headline-level annotations
annotations = read_csv(
  'headline_level_data.csv',
  show_col_types = FALSE
  )

# raw model villain and victim probabilities
villain_probs = read_csv(
  'olmo32b_villain_qlora/test_probs.csv',
  show_col_types = FALSE
  )

victim_probs  = read_csv(
  'olmo32b_victim_qlora/test_probs.csv',
  show_col_types = FALSE
  )

test_manifest = manifest %>% filter(split == "test")
cat("Test split:", nrow(test_manifest), "headlines\n")
```

## Verify Row Alignment

The test probability files do not contain ids, so bind by row order

```{r}
test_df = test_manifest %>%
  mutate(
    p_villain = villain_probs$prob_1,
    y_villain_check = villain_probs$label,
    p_victim = victim_probs$prob_1,
    y_victim_check = victim_probs$label
  )

# labels recorded alongside the probabilities must match the manifest
stopifnot(all(test_df$villain_presence_majority == test_df$y_villain_check))
stopifnot(all(test_df$victim_presence_majority == test_df$y_victim_check))

# the headline inside each prompt must match the manifest headline
prompt_headline = str_match(villain_probs$prompt,
                            'Here is the headline:\\n"(.*)"\\n')[, 2]
clean = function(x) str_squish(str_replace_all(x, "\\n", " "))
stopifnot(all(clean(prompt_headline) == clean(test_df$share_title)))

cat("Alignment verified on", nrow(test_df), "rows\n")
```

## Attach Source Quality

```{r}
test_df = test_df %>%
  left_join(annotations %>%
              distinct(url_rid, pc1, villain_presence_sum, victim_presence_sum,
                       villain_presence_count, victim_presence_count),
            by = "url_rid")
```

# Calibration Function

One function, applied to each task.

```{r}
eps = 1e-4

fit_calibration = function(df, y_col, p_col, task_label, v_folds = 10) {
  cal_df = df %>%
    transmute(
      y = as.integer(.data[[y_col]]),
      p = pmin(pmax(as.numeric(.data[[p_col]]), eps), 1 - eps),
      lp = qlogis(p),               # logit-scale predictor
      pc1 = as.numeric(pc1)
    ) %>%
    filter(!is.na(y), !is.na(p), !is.na(pc1))

  # cross-fit diagnostics: does calibration remove the error-pc1 slope?
  set.seed(42)
  folds = vfold_cv(cal_df, v = v_folds, strata = y)
  cf = folds %>%
    mutate(
      fit = map(splits, ~ glm(y ~ lp * pc1, family = binomial(),
                              data = analysis(.x))),
      preds = map2(splits, fit, ~ assessment(.x) %>%
                     mutate(p_cal = predict(.y, newdata = assessment(.x),
                                            type = "response")))
    ) %>%
    pull(preds) %>%
    bind_rows() %>%
    mutate(err_raw = p - y, err_cal = p_cal - y)

  slope_raw = tidy(lm(err_raw ~ pc1, data = cf)) %>% filter(term == "pc1")
  slope_cal = tidy(lm(err_cal ~ pc1, data = cf)) %>% filter(term == "pc1")

  diag_tbl = bind_rows(
    slope_raw %>% mutate(model = "Raw"),
    slope_cal %>% mutate(model = "Cross-fit calibrated")
  ) %>% select(model, estimate, std.error, p.value)

  diag_plot = cf %>%
    select(pc1, err_raw, err_cal) %>%
    pivot_longer(-pc1, names_to = "type", values_to = "err") %>%
    mutate(type = recode(type, err_raw = "Raw error",
                         err_cal = "Cross-fit calibrated error")) %>%
    ggplot(aes(pc1, err)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_smooth(method = "loess", span = 0.7, se = TRUE) +
    facet_wrap(~type, ncol = 1) +
    theme_classic() +
    labs(x = "Source quality", y = "Error (p - y)", title = task_label)

  # final model on ALL test-split rows
  final_fit = glm(y ~ lp * pc1, family = binomial(), data = cal_df)

  # bootstrap: stability of the calibration mapping
  set.seed(42)
  boot_curves = map_dfr(1:500, function(b) {
    bs = cal_df[sample(nrow(cal_df), replace = TRUE), ]
    fit = glm(y ~ lp * pc1, family = binomial(), data = bs)
    expand_grid(p = c(0.1, 0.5, 0.9), pc1 = c(0.2, 0.5, 0.8)) %>%
      mutate(lp = qlogis(p),
             p_cal = predict(fit, newdata = cur_data(), type = "response"),
             rep = b)
  })
  boot_tbl = boot_curves %>%
    group_by(p, pc1) %>%
    summarise(p_cal_med = median(p_cal),
              p_cal_lo = quantile(p_cal, .025),
              p_cal_hi = quantile(p_cal, .975), .groups = "drop")

  list(fit = final_fit, diagnostics = diag_tbl, plot = diag_plot,
       boot = boot_tbl, n = nrow(cal_df))
}
```

# Villain Calibration

```{r}
cal_villain = fit_calibration(test_df, "villain_presence_majority",
                              "p_villain", "Villain")
cat("n =", cal_villain$n, "\n")
kable(cal_villain$diagnostics, digits = 4)
cal_villain$plot
```

```{r}
kable(tidy(cal_villain$fit, conf.int = TRUE), digits = 3)
kable(cal_villain$boot, digits = 3,
      caption = "Bootstrap stability: calibrated probability at reference (p, pc1) points")
```

# Victim Calibration

```{r}
cal_victim = fit_calibration(test_df, "victim_presence_majority",
                             "p_victim", "Victim")
cat("n =", cal_victim$n, "\n")
kable(cal_victim$diagnostics, digits = 4)
cal_victim$plot
```

```{r}
kable(tidy(cal_victim$fit, conf.int = TRUE), digits = 3)
kable(cal_victim$boot, digits = 3)
```

# Save

```{r}
saveRDS(cal_villain$fit, "villain_probability_calibration.rds")
saveRDS(cal_victim$fit,  "victim_probability_calibration.rds")
```

# Session Info

```{r}
sessionInfo()
```
